---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dcmodifydt

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/dcmodifydt)](https://CRAN.R-project.org/package=dcmodifydt)
  [![R-CMD-check](https://github.com/data-cleaning/dcmodifydt/workflows/R-CMD-check/badge.svg)](https://github.com/data-cleaning/dcmodifydt/actions)
[![Codecov test coverage](https://codecov.io/gh/data-cleaning/dcmodifydt/branch/main/graph/badge.svg)](https://codecov.io/gh/data-cleaning/dcmodifydt?branch=main)
<!-- badges: end -->

`dcmodifydt` executes [`dcmodify`](https://CRAN.R-project.org/package=dcmodify) modification rules on a data.table, making it easy to switch between data.frame
, database or data.table implementation.

`dcmodify` separates **intent** from **execution**: a user specifies _what_, _why_ and _how_ of an automatic data change and uses dcmodify to  execute them on a `data.table`.

### Rationale

`data.table` is an excellent package with great performance. Why not code
modification rules directly in R script? _Keep it simple_ is certainly 
a good advise, reasons to use `dcmodify` are:

- You have a set "correction" rules in a production process that you apply repeatly, e.g. in each production run.

- You want to share these rules with different production processes. 

- You have data matter specialists that do not know (or have to) program in R,
and use this as a quality frame work to communicate and specify the step to 
correct or derive variables.

In these cases it is useful to label, describe and document your rules, and 
use `dcmodify`, `dcmodifydt` or `dcmodifydb` to apply the modification rules.


## Installation

The development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("data-cleaning/dcmodifydt")
```
## Example

```{r}
library(dcmodifydt)
```


```{r, code = readLines("example/modify.R")}
```

### Documented modification rules:

 Create a modifier object:
```{r}
m <- modifier(.file = "example/example.yml")
```

From this specification file "example/example.yml":

```yml
```{r, child="example/example.yml", results="asis"}
```
```

```{r}
dat <- data.table::fread(text = 
"age, income
  11,   2000
 150,    300
  25,   2000
 -10,   2000")

modify(dat, m, copy=FALSE)
print(dat)
```

For convenience it is possible to dump the data.table statements. This 
file can be sourced, resulting in the modifications of the supplied table (name)

```{r, eval = FALSE}
dump_dt(m, name = "my_dt", file="change_my_dt.R")
```

```r
```{r, echo = FALSE, results="asis"}
dump_dt(m, name = "my_dt")
```
```
